Bayesian 5 taxa family

### 2021-03-10 scripts/45_16s_sorting_out.Rmd
slowly but making progress
slowly but getting things going ...

mptb_ineli <-  read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/MSK/patients/msk_essential_data_v4.xlsx', sheet = 2)
mptb <-  read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/MSK/patients/msk_essential_data_v4.xlsx', sheet = 1) %>% 
  mutate(mrn = as.numeric(mrn)) %>% 
  filter(! mrn %in% mptb_ineli$pt_id)

@@

```{r}
m_meta_b <- mstb %>% 
  filter(str_detect(time_group, '^Baseline')) %>% 
  inner_join(mptb, by = "mrn")

@@

```{r}
m <- m_meta_b %>% 
  select(-pt_id_old) %>% 
  rename(pt_id = mrn) %>% 
  select(-last_followup) %>% 
  mutate(center = 'M') %>% 
  select(-cr_overall) # this column is now gone

@@

combined_join %>% 
  write_csv('../data/amplicon/stool/combined_2_meta.csv')
```

>>>>

mptb_ineli <-  read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/MSK/patients/msk_essential_data_v4.xlsx', sheet = 2)
mptb <-  read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/MSK/patients/msk_essential_data_v4.xlsx', sheet = 1) %>% 
  filter(! pt_id %in% mptb_ineli$pt_id)

@@

```{r}
m_meta_b <- mstb %>% 
  filter(str_detect(time_group, '^Baseline')) %>% 
  inner_join(mptb, by = c('mrn' = 'pt_id'))

@@

```{r}
m <- m_meta_b %>% 
  select(-pt_id_old) %>% 
  rename(pt_id = mrn) %>% 
  select(-last_followup) %>% 
  mutate(center = 'M') %>% 
  select(pt_id:vital_status, center) # there are several more columns in the msk table now

@@

combined_join %>% 
  write_csv('../data/amplicon/stool/combined_2_meta.csv')

combined_join %>% 
  write_csv('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/Angel_amplicon/16s_data_pheno_comprehensive.csv')
```

### 2021-03-10 scripts/53_bayesian_5fam.Rmd
slowly but getting things going ...

```{r}
# the model 
tox <- brm( toxicity  ~ 
              Akkermansia +
              Bacteroides +
              Enterococcus +
              Faecalibacterium +
              Ruminococcus +
               (1 | pt_id) ,  
              data = dat, 
              family = bernoulli(), 
              warmup = 5000, iter = 40000, 
              cores = 8, chains = 2, 
              seed = 123)

@@

      res = posterior_samples(mod) %>% 
              select(starts_with('b')) %>% 
              select(-b_Intercept)
      return(res)
  }) %>% 
  bind_cols()

```

>>>>

```{r}
# the model 
# tox <- brm( toxicity  ~ 
#               Akkermansia +
#               Bacteroides +
#               Enterococcus +
#               Faecalibacterium +
#               Ruminococcus +
#                (1 | pt_id) ,  
#               data = dat, 
#               family = bernoulli(), 
#               warmup = 5000, iter = 40000, 
#               cores = 8, chains = 2, 
#               seed = 123)

@@

      res = posterior_samples(mod) %>% 
              select(starts_with('b')) %>% 
              select(-b_Intercept)
      return(res)
  }) %>%  
  bind_cols() %>% 
  gather('genus','coeff')
```

```{r}
# plotting in ridge plot
df <- ret %>% 
  group_by(genus) %>% 
  mutate(meanpergenus = mean(coeff))%>%
  ungroup()  %>% 
  mutate(genus = str_replace(genus, 'b_',''))

ggplot()+
  ggridges::geom_density_ridges(data  = df, 
                                aes(x      = coeff,
                                    y      = reorder(as.factor(genus), meanpergenus),
                                    height = ..density..),
                                scale = 1, size = 0.25,
                                alpha = 0.6) +
  geom_vline(xintercept = 0, col = "red", linetype = 'dashed', size = 1) +
  #scale_x_continuous(limits = c(-0.02,0.03)) +
  labs(y = '',
       title = 'Response: toxicity') +
  ggthemes::theme_tufte() +
  theme(axis.text.y = element_text(size=20),
        axis.text.x = element_text(size=15)) +
  ggsave('../figs/bayesian_genus_5_each_toxicity.jpg', width = 8)    
```

```{r}
# also when the response is cr_d100
ret_CR <- target %>% 
  set_names(target) %>% 
  map(function(genus) {
      mod =  brm( as.formula(str_glue('cr_d100  ~ 
              {genus} +
               (1 | pt_id)')),  
             data = dat_cr, 
              family = bernoulli(), 
              warmup = 3000, iter = 10000, 
              cores = 8, chains = 2, 
              seed = 123)

      res = posterior_samples(mod) %>% 
              select(starts_with('b')) %>% 
              select(-b_Intercept)
      return(res)
  }) %>%  
  bind_cols() %>% 
  gather('genus','coeff')
```
```{r}
ret_CR %>% 
  group_by(genus) %>% 
  mutate(meanpergenus = mean(coeff))%>%
  ungroup()  %>% 
  mutate(genus = str_replace(genus, 'b_','')) %>% 
  ggplot()+
  ggridges::geom_density_ridges(aes(x      = coeff,
                                    y      = reorder(as.factor(genus), meanpergenus),
                                    height = ..density..),
                                scale = 1, size = 0.25,
                                alpha = 0.6) +
  geom_vline(xintercept = 0, col = "red", linetype = 'dashed', size = 1) +
  #scale_x_continuous(limits = c(-0.02,0.03)) +
  labs(y = '',
       title = 'Response: cr_d100') +
  ggthemes::theme_tufte() +
  theme(axis.text.y = element_text(size=20),
        axis.text.x = element_text(size=15)) +
  ggsave('../figs/bayesian_genus_5_each_CR.jpg', width = 8)   
```
